Capstone Movielens Report

### 2025-09-23 reports/capstone-movielens.site/02-0.methods.Rmd
Fix for Section 2.2 Preparing train and set datasets.

::: { #func.make_source_datasets .sidebar }
  - [make_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L87)
:::

::: { #func.init_source_datasets .sidebar }
  - [init_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L209)
:::

@@

### The [make_source_datasets](#func.make_source_datasets) function
\

@@

make_source_datasets <- function(){
  # ...
  list(edx_CV = edx_CV,
       edx.mx = edx.mx,
       edx.sgr = edx.sgr,
       tuning_sets = tuning_sets,
       movie_map = movie_map,
       date_days_map = date_days_map)
}
```

>>>>

  - [init_source_datasets](#func.init_source_datasets)

  - [make_source_datasets](#func.make_source_datasets)

@@

### [init_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L209) Function {#func.init_source_datasets}

The _init_source_datasets_ function checks whether the required source datasets are cached on the hard disk, and, if so, loads the datasets from the cache file.

Otherwise, it calls the [make_source_datasets](#func.make_source_datasets) function, which creates all the necessary datasets.

Finally, it stores the data into a cache file on the disk.

The simplified version of the function's source code is shown below:

```{r eval=FALSE}
init_source_datasets <- function(){
  put_log("Method `init_source_datasets`: 
Initializing sourse datasets...")

  if(file.exists(movielens_datasets_file_path)){
    movielens_datasets <- load_movielens_data_from_file(movielens_datasets_file_path)
  } else if(file.exists(movielens_datasets_zip)) {
    put_log("Method `init_source_datasets`: 
Unzipping MovieLens data file from zip-archive: {movielens_datasets_zip}...")

    unzip(movielens_datasets_zip, movielens_datasets_file_path)

    if(!file.exists(movielens_datasets_file_path)) {
      put_log("Method `init_source_datasets`: 
File does not exists: {movielens_datasets_file}.")
      stop("Failed to unzip MovieLens data zip-archive.")
    }

    movielens_datasets <- load_movielens_data_from_file(movielens_datasets_file_path)
  } else {
    put_log("Method `init_source_datasets`: 
Creating datasets...")
    library(edx.capstone.movielens.data)
    put_log("Method `init_source_datasets`: 
Library attached: 'edx.capstone.movielens.data'")

    movielens_datasets <- make_source_datasets()
    put("Method `init_source_datasets`: 
All required datasets have been created.")

    put_log("Method `init_source_datasets`: 
Saving newly created input datasets to file...")
    save(movielens_datasets, file =  movielens_datasets_file_path)

    if(!file.exists(movielens_datasets_file_path)) {
      put_log("Method `init_source_datasets`: 
File was not created: {movielens_datasets_file}.")
      warning("MovieLens data was not saved to file.")
    } else {
      put_log("Method `init_source_datasets`: 
Datasets have been saved to file: {movielens_datasets_file_path}.")

      put_log("Method `init_source_datasets`: 
Creating zip-archive: {movielens_datasets_zip}...") 
      zip(movielens_datasets_zip, movielens_datasets_file_path)
    # ...
    }
  }
  movielens_datasets
}
```

### [make_source_datasets](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L87) Function {#func.make_source_datasets}

@@

```{r eval=FALSE}
make_source_datasets <- function(){
  # ...
  list(edx_CV = edx_CV,
       edx.mx = edx.mx,
       edx.sgr = edx.sgr,
       tuning_sets = tuning_sets,
       movie_map = movie_map,
       date_days_map = date_days_map)
}
```
